%% main.tex
%% Is the AI Infrastructure Boom an Economic Bubble?
%% Quantitative Evidence from Hyperscaler Capex-FCF Divergence,
%% IPO Supply Shocks, and Index Concentration Risk
%%
%% Submission target: arXiv q-fin.RM / q-fin.ST
%% Document class: standard article with amsmath

\documentclass[11pt,letterpaper]{article}

%% ── Packages ────────────────────────────────────────────────────────────────
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}

%% Mathematics
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bm}

%% Figures, Tables
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{longtable}
\usepackage{float}

%% Formatting
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{setspace}
\setstretch{1.15}
\usepackage{parskip}

%% Code listings
\usepackage{listings}
\lstset{basicstyle=\small\ttfamily, breaklines=true}

%% Colours
\usepackage[dvipsnames]{xcolor}
\definecolor{linkblue}{RGB}{0, 70, 127}

%% Hyperlinks (arXiv-compatible)
\usepackage[
  colorlinks = true,
  linkcolor  = linkblue,
  citecolor  = linkblue,
  urlcolor   = linkblue,
  pdftitle   = {Is the AI Infrastructure Boom an Economic Bubble?},
  pdfauthor  = {Vinesh Ranga},
]{hyperref}

%% Bibliography
\usepackage[numbers,sort&compress]{natbib}
\bibliographystyle{plainnat}

%% Caption formatting
\usepackage[font=small,labelfont=bf,skip=4pt]{caption}

%% ── Custom Commands ──────────────────────────────────────────────────────────
\newcommand{\VaR}{\text{VaR}}
\newcommand{\PFE}{\text{PFE}}
\newcommand{\HHI}{\text{HHI}}
\newcommand{\MLB}{\text{MLB}}
\newcommand{\betaliq}{\beta_{\text{liq}}}
\newcommand{\zCF}{z_{\text{CF}}}

%% ── Document ─────────────────────────────────────────────────────────────────
\begin{document}

\title{
  \textbf{Is the AI Infrastructure Boom an Economic Bubble?}\\[6pt]
  \large Quantitative Evidence from Hyperscaler Capex-FCF Divergence,\\
  IPO Supply Shocks, and Index Concentration Risk
}
\author{
  Vinesh Ranganathan\\[4pt]
  \small \texttt{https://github.com/VineshRanga/Is-AIBubble}
}
\date{June 2026 \quad (Working Paper)}

\maketitle

\begin{abstract}
The four largest AI hyperscalers---Microsoft, Alphabet, Amazon, and Meta---deployed a
combined \$293 billion in capital expenditure during fiscal year 2025, generating only
\$127 billion in free cash flow. The resulting Capex-to-FCF ratio of $2.31\times$ crossed
the structural stress threshold of $1.0\times$ in the first quarter of 2026.
We term this divergence the \textit{Compute Cliff}.

Simultaneously, three pure-play AI entities (SpaceX, Anthropic, and OpenAI) are targeting
public market listings that would collectively drain an estimated \$210 billion in
institutional liquidity within a six-month window (June--December 2026). This supply shock
arrives at the worst possible macro moment: when hyperscaler balance sheets are under maximum
stress, and when the Nasdaq-100 is at its highest historical concentration level
(top-5 constituents represent approximately 68\% of index weight).

We build a quantitative risk framework to evaluate three probability paths for public equity
markets through December 2027: a \textit{Fizzle} scenario (mild $-12.5\%$ compression
absorbed gracefully), a \textit{Systemic Short} scenario ($-30\%$ over 90 trading days
driven by macro hedge fund amplification), and a \textit{Dotcom Pop} trajectory ($-55\%$
over 18 months replicating the 2000--2002 structural liquidation velocity).

Our Index Liquidity Elasticity Coefficient $\betaliq$, calibrated on the 2022 tech drawdown
as an out-of-sample anchor, produces a 99th-percentile Potential Future Exposure (PFE) of
$75.5\%$ for the Nasdaq-100 under the Dotcom trajectory. A formal Kupiec (1995) and
Christoffersen (1998) backtesting framework confirms that the Cornish-Fisher VaR model
is appropriately conservative for the 2022--2026 evaluation window.

The paper concludes with a discussion of the limitations inherent in this framework,
including the use of public market proxies for private foundational lab valuations, the
assumption of flat FCF baselines, and the potential for compute derivatives to alter
the modelled correlation structure.
\end{abstract}

\tableofcontents
\newpage

%% ── Section 1: Introduction ──────────────────────────────────────────────────
\section{Introduction}
\label{sec:intro}
\input{sections/01_catalyst}

%% ── Section 2: Module 2 Impact (Concentration, VaR, PFE) ────────────────────
\section{Market Impact: Index Concentration and Tail Risk}
\label{sec:impact}
\input{sections/02_impact}

%% ── Section 3: Mitigation and Policy Implications ────────────────────────────
\section{Risk Mitigation and Policy Implications}
\label{sec:mitigation}
\input{sections/03_mitigation}

%% ── Section 4: Limitations and Future Work ───────────────────────────────────
\section{Methodology Limitations and Future Work}
\label{sec:limitations}
\input{sections/04_limitations}

%% ── Section 5: Conclusion ────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}

The AI infrastructure buildout represents the most concentrated capital deployment cycle
in the history of the technology sector, executed at speeds that dwarf the 1990s
telecommunications dark-fibre boom. The Compute Cliff---a structural divergence between
capital deployment velocity and monetisation capacity---creates a precondition for
balance-sheet stress at the hyperscaler layer even before external market shocks are
considered.

When the $\$210$B IPO liquidity vacuum is introduced into this already-stressed system,
the Systemic Short scenario ($-30\%$ NDX compression within 90 trading days) emerges
as the most historically consistent outcome, replicating the 2022 velocity profile under
a supply-driven rather than rate-driven trigger. The elevated index concentration
(NDX HHI: 874 in 2026 vs.\ 297 in 2010; a $2.78\times$ amplifier) ensures that any
macro shock to the AI cluster propagates with non-linear force through passive vehicle
holders.

Future work should instrument actual compute derivatives pricing, incorporate real-time
index weight data from paid providers, and extend the PFE framework to second-order
credit and volatility surface effects.

%% ── Bibliography ─────────────────────────────────────────────────────────────
\bibliography{references}

%% ── Appendices ───────────────────────────────────────────────────────────────
\appendix

\section{Data Sources and Reproducibility}
\label{app:data}

All data used in this study is sourced from publicly available, open-source endpoints
and is fully reproducible:

\begin{itemize}
  \item \textbf{Hyperscaler financials:} SEC EDGAR CompanyFacts XBRL API
    \citep{sec_xbrl_api}, queried via the project's \texttt{src/etl/edgar.py} module.
    CIKs: MSFT 0000789019 \citep{msft_10k_2025}, GOOGL 0001652044 \citep{googl_10k_2025},
    AMZN 0001018724 \citep{amzn_10k_2025}, META 0001326801 \citep{meta_10k_2025}.
  \item \textbf{Market data:} Yahoo Finance via \texttt{yfinance} \citep{yfinance}.
    Tickers: \texttt{\^{}GSPC}, \texttt{\^{}NDX}, MSFT, GOOGL, AMZN, META, CSCO.
  \item \textbf{Macro indicators:} Federal Reserve FRED API via \texttt{pandas-datareader}.
    Series: FEDFUNDS \citep{fred_fedfunds}, M2SL \citep{fred_m2},
    STLFSI4 \citep{fred_stlfsi}.
  \item \textbf{Code repository:} \url{https://github.com/VineshRanga/Is-AIBubble}
\end{itemize}

\section{Cornish-Fisher VaR Derivation}
\label{app:cf}

The Cornish-Fisher modification adjusts the standard normal quantile $z_\alpha$ for
observed third and fourth cumulants (skewness $\hat{S}$ and excess kurtosis $\hat{K}$):

\begin{equation}
  \zCF = z_\alpha
       + \frac{z_\alpha^2 - 1}{6}\hat{S}
       + \frac{z_\alpha^3 - 3z_\alpha}{24}\hat{K}
       - \frac{2z_\alpha^3 - 5z_\alpha}{36}\hat{S}^2
  \label{eq:cf_expansion}
\end{equation}

The Cornish-Fisher VaR is then:

\begin{equation}
  \VaR_{\text{CF}} = -\hat{\mu} + \hat{\sigma} \cdot \zCF
  \label{eq:cf_var}
\end{equation}

where $\hat{\mu}$ and $\hat{\sigma}$ are the sample mean and standard deviation of the
return distribution estimated on the in-sample calibration window.

\section{Kupiec and Christoffersen Test Statistics}
\label{app:tests}

\subsection{Kupiec Proportion-of-Failures Test}
\label{app:kupiec}

Let $T$ be the number of OOS observations and $x$ the observed number of VaR breaches.
Under $H_0$: $p_{\hat{\text{}} } = x/T = p$ (theoretical breach probability), the
log-likelihood ratio statistic is:

\begin{equation}
  LR_{\text{POF}} = -2\ln\!\left[\frac{p^x(1-p)^{T-x}}{\hat{p}^x(1-\hat{p})^{T-x}}\right]
  \sim \chi^2(1)
  \label{eq:kupiec_lr}
\end{equation}

\subsection{Christoffersen Independence Test}
\label{app:christoffersen}

Let $I_t \in \{0,1\}$ be the breach indicator at time $t$. Define transition probabilities
$\pi_{01} = P(I_t = 1 \mid I_{t-1} = 0)$ and $\pi_{11} = P(I_t = 1 \mid I_{t-1} = 1)$.
Under $H_0$ (independence): $\pi_{01} = \pi_{11} = \pi_2$ (unconditional). The LR statistic is:

\begin{equation}
  LR_{\text{ind}} = 2\ln\!\left[
    \frac{(1-\hat\pi_{01})^{n_{00}} \hat\pi_{01}^{n_{01}}
          (1-\hat\pi_{11})^{n_{10}} \hat\pi_{11}^{n_{11}}}
         {(1-\hat\pi_2)^{n_{00}+n_{10}} \hat\pi_2^{n_{01}+n_{11}}}
  \right] \sim \chi^2(1)
  \label{eq:christoffersen_lr}
\end{equation}

The combined conditional-coverage test statistic $LR_{\text{CC}} = LR_{\text{POF}} + LR_{\text{ind}} \sim \chi^2(2)$.

\end{document}
